Between Brussels and Britain: Uncovering the driving forces of Brexit

### 2024-04-18 paper/paper.qmd
Model

---
title: "Between Brussels and Britain: Understanding Party Engagement with EU and Domestic Topics in the UK Parliament"
subtitle: "A statistical approach to assessing engagement in a supranational union"
author: 
  - Samarth Rajani
thanks: "Code and data are available at: https://github.com/samarthrajani/-Between-Brussels-and-Britain-Understanding-Party-Engagement-with-EU-and-Domestic-Topics-in-the-UK-."
date: today
date-format: long
abstract: "In a union, all parties must work together to guarantee best outcomes. In the case of the United Kingdom, friction between legislative bodies within the nation and the European Parliament caused the UK to exit the union. This paper looks at a speech based dataset from the UK House of Commons to suggest which political parties were more vocal in the build up to Brexit, giving insight into which political parties were the driving force in this change.  "
format: pdf
number-sections: true
bibliography: references.bib
---

@@

@@

@@

Define $y_i$ as discussion of EU in that sentence. Then $\beta_i$ is the party affiliation.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

#first_model <-
  #readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

#modelsummary::modelsummary(
  #list(
#    "First model" = first_model
 # ),
  #statistic = "mad",
  #fmt = 2
#)
```

@@

The one obvious limitation of this paper is that it only uses data from 2014 to 2016 in its analysis to judge which parties are being vocal. This was done due to resource limitations in the logistic analysis - the runtime for merely 2 years was close to 5 hours, severely limiting the ability to measure the relationship between party affiliation and talking about the EU in the House of Commons. Since the Brexit referendum took place in 2016, it is possible that the propensity to talk about the EU was in fact talking just about the referendum, and therefore might not be the vocalisation of disagreement. This causes our estimates to overestimate the the friction between the political parties and European Parliament, making our conclusions inaccurate.

@@

In the release note, the authors discuss how a binary classification of EU related issues may have been too strict, and thereby generated more false negatives (classifier incorrectly classifies a sentence discussing EU Governance as not discussing EU governance) in the pursuit of less false positives (classifier incorrectly classifies a sentence as discussing EU Governance when it does not). This causes a potentially downward bias in our logistic regression estimates, underestimating the extent to which EU related matters have dictated discourse

## Next steps

Possibly construct a

>>>>

---
title: "Between Brussels and Britain: Uncovering the driving forces of Brexit"
subtitle: "A statistical approach to assessing engagement in a supranational union"
author: 
  - Samarth Rajani
thanks: "Code and data are available at: https://github.com/samarthrajani/-Between-Brussels-and-Britain-Understanding-Party-Engagement-with-EU-and-Domestic-Topics-in-the-UK-."
date: today
date-format: long
abstract: "In a union, all parties must work together to guarantee best outcomes. In the case of the United Kingdom, friction between legislative bodies within the nation and the European Parliament caused the UK to exit the union. This paper looks at a speech based dataset from the UK House of Commons to suggest which political parties were more vocal in the build up to Brexit, giving insight into which political parties were the driving force in this change.  "
format: pdf
number-sections: true
bibliography: references.bib
---

@@

## Variables

@@

As can be seen, there is very little variation in discussion of specific policy areas over the course of 10 years. This could be due to the dataset being extremely large, as it would iron out any major differences due to sheer sample size. There is also a case to be made that the CAP classifier used by the authors to assign the policy area to sentences may not be completely accurate, although that discussion is not relevant to this paper.

@@

Define $y_i$ as discussion of the EU in a specific sentence of a speech. Then $x_i$ is the party affiliation.

```{=tex}
\begin{align*} 
y_i|\pi_i &\sim \text{Bernoulli}(\pi_i) \\
\text{logit}(\pi_i) &= \beta_0 + \beta_1 x_i + \gamma_i\\
\beta_0 &\sim \text{Normal}(0, 2.5) \\
\beta_1 &\sim \text{Normal}(0, 2.5) 
\end{align*}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Logistic model of likelihood of a party affiliation speaking about the EU"
#| warning: false

summary(first_model)
```

@@

The one obvious limitation of this paper is that it only uses data from 2014 to 2016 in its analysis to judge which parties are being vocal. This was done due to resource limitations in the logistic analysis - the runtime for merely 2 years was close to 5 hours, severely limiting the ability to measure the relationship between party affiliation and talking about the EU in the House of Commons. Since the Brexit referendum took place in 2016, it is possible that the propensity to talk about the EU was in fact talking just about the referendum, and therefore might not be the vocalisation of disagreement. This causes our estimates to overestimate the the friction between the political parties and European Parliament, making our conclusions inaccurate. This is a fundamental reverse causality problem.

@@

In the release note, the authors discuss how a binary classification of EU related issues may have been too strict, and thereby generated more false negatives (classifier incorrectly classifies a sentence discussing EU Governance as not discussing EU governance) in the pursuit of less false positives (classifier incorrectly classifies a sentence as discussing EU Governance when it does not). This causes a potentially downward bias in our logistic regression estimates, underestimating the extent to which EU related matters have dictated discourse.

## Next steps - An Instrumental Variable Framework

This paper lays the groundwork for a more detailed study into which parties played the strongest role in Brexit. With the dataset of speeches already avaiable, devising an Intrumental Variable Framework to determine party influence seems like a suitable way forward.

As previously discussed, since speech occurence may not be a robust proxy for disagreement, employing an Instrumental Variable strategy will get rid of the reverse causality concerns.

### Instrumental Variable

When studying the relationship between two variables, there is a concern of reverse causality. For example, if we were to study the impact on institutions on the GDP of a nation,
